An integrated distance sampling approach for modelling age-structured populations

### 2023-08-02 Manuscript_Nilsen_Nater.qmd
Some text updates

Estimating abundance and demographic rates for wildlife populations is a key element of population ecology, demography, eco-evolution research, conservation and wildlife management (refs). Tremendous progress during the last few decades both in terms of new technologies that allow new types of field data being collected (e.g. eDNA ref; camera traps; refs) and the statistical machinery available to interpret the ecological and demographic signals in the data (refs). Nevertheless, large scale and long term data sets are still relatively rare (ref), due to high costs of field data collection.

>>>>

Estimating abundance and demographic rates for wildlife populations is an integral part of basic and applied ecology [@Skalski2005]. Over the last few decades, a tremendous progress have been made towards this end. This progress is partly driven the development and application of new field data collection methods and apporaches, such as citizen science data (ref), camera trap data (ref) and the collection of environmental DNA data (ref). In addition, developments of new analysis methods allow us to estimate abundance and demographic rates it situations where it was not feasible before (ref). combined, these advance put us in a much better position to estimate state variables needed for population management (ref) but also estimate indexes that are relevant for large scale policy applications, e.g. such as Essential Biodiversity Variables (ref).

Nevertheless, large scale and long term data sets are still relatively rare (ref), due to high costs of field data collection.
